Local to global N input modelling

### 2025-05-16 08_manuscript/08_manuscript.qmd
New output structure in manuscript

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(here)
library(ggpubr)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()
```

@@

@@

* Farms with mean temperature < 4$^\circ$ were excluded

@@

pop_params <- tar_read(pop_params, store = "../05_targets_farm")
species_params <- tar_read(species_params, store = "../05_targets_farm")
params <- c(pop_params, species_params)
pnames <- names(params)
sens_excr <- file.path("..", "data", "atlantic_salmon", "data_products", "excreted_parameter_sensitivity.parquet") %>%
  read_parquet() %>%
  mutate(adj_param = factor(adj_param, levels = rev(adj_param)))
```

@@

farm_harvest_size <- tar_read(farm_harvest_size, store = "../04_targets_individual")

ggplot(farm_harvest_size, aes(x = weight/1000)) + 
  geom_histogram(colour = "black", fill = "salmon", alpha = 0.5) + 
  theme_classic() + 
  scale_x_continuous(breaks = seq(0, 8, 1), limits = c(1.25, 7.75)) + 
  labs(y = "Frequency", x = "Individual weight (kg)")
```

>>>>

library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggh4x)
library(here)
library(ggpubr)
library(patchwork)
library(qs)
library(arrow)
library(targets)
library(sf)
library(terra)
library(rnaturalearth)
library(rnaturalearthdata)
library(dtplyr)
library(furrr)
library(RColorBrewer)
library(purrr)
library(future)
library(readxl)
library(units)
library(conflicted)
library(lubridate)
conflicts_prefer(dplyr::select(), dplyr::filter(), dplyr::intersect(), .quiet = T)
# This markdown uses TinyTex - install with tinytex::install_tinytex()
```

```{r global-variables}
#| include: false
#| eval: true

source(here("00_model_functions.R"))

# Input paths
input_farm_coords_path <- here() %>% file.path("data", "_general_data", "farm_locations")
input_farm_sst_path <- here() %>% file.path("data", "_general_data", "SST")
input_species_param_path <- here() %>% file.path("data", "atlantic_salmon", "params")
input_feed_profile_path <- here() %>% file.path("data", "_general_data", "diets")

# Output paths
output_farm_data_path <- here() %>% file.path("outputs", "farm_data")
output_species_data_path <- here() %>% file.path("outputs", "species_data")
output_sens_data_path <- here() %>% file.path("outputs", "sensitivity_data")
output_farm_growth_data_path <- here() %>% file.path("outputs", "farm_growth_data")
output_cohort_growth_data_path <- here() %>% file.path("outputs", "cohort_growth_data")

# Filenames
farm_coords_file <- file.path(output_farm_data_path, "farm_coords.qs")
species_params_file <- file.path(output_species_data_path, "species_params.qs")
pop_params_file <- file.path(output_species_data_path, "pop_params.qs")
feed_params_file <- file.path(output_species_data_path, "feed_params.qs")
sens_params_file <- file.path(output_species_data_path, "sens_params.qs")
sens_results_files <- file.path(output_sens_data_path) %>% str_subset("results")

species_params <- qread(species_params_file)
pop_params <- qread(pop_params_file)
```

@@

```{r farm temperature data}
farms_to_omit <- file.path(input_farm_coords_path, "atlantic_salmon_farms_to_omit.qs") %>% qs::qread()
farm_ts_data <- file.path(output_farm_data_path, "farm_ts_data.qs") %>% qs::qread()
farm_coords <- file.path(output_farm_data_path, "farm_coords.qs") %>% qs::qread()
farms_geometry <- file.path(output_farm_data_path, "farm_geometry.qs") %>% qs::qread() %>% 
  rename(farm_ID = farm_id)

farm_temp_means <- farm_ts_data %>% 
  group_by(farm_ID) %>% 
  reframe(mean_temp = mean(temp_c)) %>% 
  merge(farms_geometry, by = "farm_ID") %>% 
  merge(farm_coords, by = "farm_ID") %>% 
  dplyr::filter(!farm_ID %in% farms_to_omit)

overall_mean <- mean(farm_temp_means$mean_temp)
overall_sd <- sd(farm_temp_means$mean_temp)
```

```{r fig-farm-mean-temps}
#| fig-caption: Colours show farm site mean temperature

worldmap <- ne_countries(scale = "medium", returnclass = "sf")
farms_to_omit <- here() %>% 
  file.path("data", "_general_data", "farm_locations", "atlantic_salmon_farms_to_omit.qs") %>% 
  qread()

farms_w_temp_df <- here() %>% 
  file.path("data", "_general_data", "farm_locations", "atlantic_salmon_locations_w_temps.qs") %>% 
  qread() %>% 
  filter(!farm_id %in% farms_to_omit) %>% 
  mutate(country = as.factor(country),
         species_group = as.factor(species_group),
         data_type = as.factor(data_type),
         data_type_2 = as.factor(data_type_2),
         model_name = as.factor(model_name),
         F_CODE = as.factor(F_CODE),
         day = str_split_i(day, "_", 2) %>% as.integer())

farm_temp_means <- farms_w_temp_df %>% 
  select(iso3c, country, farm_id, temp_c)
farm_temp_means_1 <- farms_w_temp_df %>% 
  filter(day == 1) %>% 
  select(geometry)

farm_temp_means <- farm_temp_means %>% 
  group_by(iso3c, country, farm_id) %>% 
  reframe(temp_c = meanna(temp_c))

farm_temp_means <- cbind(farm_temp_means, farm_temp_means_1)

ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = temp_c), size = 1.5) +
  coord_sf() +
  theme_classic()

  theme_classic() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

* Farms with a mean temperature of $\leq 6^\circ$C were excluded - a total of `r as.integer(length(farms_to_omit))` farms.

[Need to put the cohort concept (and harvest timing) in here.]{.aside}

```{r fig-cohorts}
#| fig-cap: Example of cohorts

fnms <- output_cohort_growth_data_path %>% 
  list.files(full.names = T) %>% 
  str_subset("biomass")

df <- fnms[1] %>% read_parquet() %>% 
  filter(feed == "reference") %>% 
  mutate(mean = set_units(mean, "g") %>% set_units("t"),
         sd = set_units(sd, "g") %>% set_units("t"))

ggplot(df, aes(x = yday, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  scale_x_continuous(breaks = seq(0, 2000, 200)) +
  labs(y = "Farm biomass", x = "Day of the year") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 0.75)

yrs <- max(df$yday)/365
harvests <- df %>% 
  group_by(cohort) %>% 
  reframe(mean = maxna(mean),
          sd = maxna(sd))

sum(harvests$mean)/yrs
mean(harvests$sd)/yrs
```

### Feeding rate

```{r fig-functional-response-to-temperature}
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dashed line), relative to maximum potential ingestion.

df <- data.frame(water_temp = seq(-5, 26, 0.25))
df$feeding <- sapply(FUN = feeding_rate, X = df$water_temp, species_params = species_params)
df <- df %>% 
mutate(ing_pot = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * feeding,
       food_prov = NA)
for (i in 1:nrow(df)) {
  df$food_prov[i] <- food_prov_rate(pop_params = pop_params, rel_feeding = df$feeding[i], ing_pot = df$ing_pot[i], 
                                    ing_pot_10 = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * 0.1)
}
df$ing_pot <- df$ing_pot/maxna(df$ing_pot)
df$food_prov <- df$food_prov/maxna(df$ing_pot)

df %>% 
  ggplot(aes(x = water_temp, y = ing_pot)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov), linewidth = 0.75, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Relative food ingested/provided") +
  theme_classic() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

@@

params <- file.path(output_species_data_path, "sens_params.qs") %>% qs::qread()

param_names <- tibble::tribble(
  ~name,        ~lab,
  "alpha",      bquote(alpha),
  "epsprot",    bquote(epsilon["P"]),
  "epslip",     bquote(epsilon["L"]),
  "epscarb",    bquote(epsilon["C"]),
  "epsO2",      bquote(epsilon["O"[2]]),
  "pk",         bquote("pk"),
  "k0",         bquote("k"[0]),
  "m",          bquote("m"),
  "n",          bquote("n"),
  "betac",      bquote(beta*"C"),
  "Tma",        bquote("T"["max"]),
  "Toa",        bquote("T"["opt"]),
  "Taa",        bquote("T"["min"]),
  "omega",      bquote(omega),
  "a",          bquote("a"),
  "k",          bquote("k"),
  "eff",        bquote("eff"),
  "meanW",      bquote(bar("W")),
  "deltaW",     bquote(Delta*"W"), 
  "meanImax",   bquote(bar("I"["max"])),
  "deltaImax",  bquote(Delta*"I"["max"]),
  "overFmean",  bquote(bar("F")),
  "overFdelta", bquote(Delta*"F"),
  "mortmyt",    bquote("mort")
)
```

@@

file.path(output_farm_data_path, "farm_harvest_size.qs") %>% 
  qread() %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg")) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.5) +
  scale_y_continuous(limits = c(0,500)) +
  labs(y = "Frequency", x = "Harvest weight") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 0.75)
```
